\chapter{Goals}
\label{sec:goals}
This chapter focuses on the goals of Assurer. We will describe the goals that should be met by the protocol and are going to be verified by means of a protocol prover. First and foremost, the protocol should ensure the clients will be authenticated to the server and vice versa. These properties should prevent susceptibility to man-in-the-middle attacks. Furthermore both the all application data sent between client and server should be secret and trustworthy. Also the protocol should be resistant against replay attacks targeting ABCs. Finally the protocol should be designed in such a way that the connection to the server will not cause a denial of service.

To summarize the protocol should offer:
\begin{itemize}
  \item authentication of client to server;
  \item authentication of server to client;
  \item integrity of passport data sent between server and client;
  \item confidentiality of passport data sent between server and client;
  \item integrity of ABCs sent between server, client and card;
  \item confidentiality of ABCs sent between server, client and card;
  \item resistance against replay attacks.
\end{itemize}

In addition, we wish to achieve perfect forward secrecy, meaning an attacker cannot derive previous session keys even if private keys are obtained (see chapter~\ref{subsubsec:pfs}). This means we have to use cryptographically strong keys, which also should be ephemeral.

Note that replay attacks are partly mitigated by policies surrounding the system. For example, replaying of ABCs is only possible should an attacker have access to the assurer's tablet, which is kept in a safe and is inaccessible. To improve upon this, implementations of the protocol should ensure it does not allow or facilitate storing ABCs internally for repeated use. All ABCs should be stored on the corresponding IRMA card and securely deleted immediately after.

\section{Attack scenarios}
This section describes possible attack scenarios with respect to the goals described above. These are attacks the protocol should be resistant against. The possible impact of such an attack is described directly after the attack itself, along with the likelihood of such an attack occurring.

\begin{itemize}
	\item An attacker reads the passport data sent between client and server. This attack causes personally identifiable information to be learned, leading to a breach in confidentiality. This may be likely in the event weak cryptography is used or if either party loses its key. However because of the requirement of perfect forward secrecy this issue is partly mitigated (see section~\ref{subsubsec:pfs}).
  \item An attacker modifies the passport data sent between client and server. This attack causes the server to receive data inconsistent with the passport data sent by the client, at worst leading to the issuance of ABCs not corresponding to the passport. This can be mitigated, like the previous scenario, with the use of sufficiently strong cryptography, but is still somewhat likely to occur with a powerful adversary.
  \item An attacker modifies signed ABCs. This attack causes the client to receive incorrect ABCs, at worst leading to the issuance of ABCs not corresponding to the passport. This is highly unlikely as it would require the attacker to be able to sign the ABCs using the issuer's private key to keep relying parties from detecting the fraud.
  \item An attacker intercepts ABCs and stores it onto his own card. This attack causes false issuance of ABCs, in turn leading to possible fraud. This may be a likely scenario if the attacker controls the network and is actively attacking the protocol. A mitigating factor is the fact the attacker must first obtain access to the tablet, which as described is kept under lock and key.
  \item An attacker submits forged passport data to the server. This attack causes the server to possibly provide ABCs for fictitous people, in turn leading to possible fraud. Because the server verifies all passport data before providing ABCs, it may also be possible to query the server, guessing correct personally identifiyable information.
  \item An attacker intercepts and later reuses ABCs already stored onto a card for storing on another card. This attack causes the attacker to be able to commit fraud. \textsc{TODO: Likely?}
  \item An attacker replays the sending of passport data to obtain another identical set of ABCs. This attack effectively facilitates the cloning of cards. Since the server is keeping a log of all of the ABCs that are given out a simple lookup would reveal the replay attack, making this an unlikely scenario.
  \item An attacker uses different ABCs from different people to authenticate. This attack allows combining ABCs to achieve the required combination for authentication. This is hardly an issue, due to the fact the IRMA card features a photo that should resemble the authenticating person. Adding to that the fact that switching cards midway would raise eyebrows leads us to think this attack is unlikely to occur.
  \item An attacker causes a denial of service on the client. This may cause clients to not receive the requested ABCs, causing them to restart the process. This does not lead to a security risk.
  \item An attacker causes a denial of service on the server. This halts all activity surrounding the issuance of ABCs, but does not lead to a security risk.
  \item An attacker shows up with someone else's passport. At worst this could cause incorrect ABCs to be placed onto an IRMA card, however this is mitigated by verifying the photograph printed on the passport. In order to pass this test the attacker has to forge the passport, which is detected upon verification by either the client or the server. This scenario therefore is unlikely.
  \item An attacker switches out the non-initialized IRMA cards before they reach the assurer's office. This also is a non-issue, because these cards only contain a unique ID at this point in the process. This ID is not used by the relying parties for verification of ABCs and therefore has no impact on the security and privacy of the system.
\end{itemize}

\subsection{Perfect Forward Secrecy}
\label{subsubsec:pfs}
In key exchange protocols there is a property called Perfect Forward Secrecy (PFS) that provides a more secure way of encryption with respect to everyday session encryption, because the key is deleted immediately after use and therefore cannot be stolen by an attacker or forced to be handed over by the government in an attempt to decrypt intercepted traffic~\cite{lecture2}. More specifically, the exposure of long-term keying material, used in the protocol to negotiate session keys, does not compromise the secrecy of session keys established before the exposure. This property is especially relevant to scenarios in which exchanged session keys require secrecy protection beyond their lifetime, such as in the case of session keys used for data encryption. This means it is relevant in Assurer, since a lot of sensitive information is being transfered and therefore needs to be kept secret.

The most common way to achieve PFS in a key-exchange protocol is by using the Diffie-Hellman key agreement with ephemeral exponents to establish the value of a session key, while confining the use of the longterm keys (such as private signature keys) to the purpose of authenticating the exchange (see authentication). One essential element for achieving PFS with the Diffie-Hellman exchange is the use of ephemeral exponents which are erased from memory as soon as the exchange is complete. This should include the erasure of any other information from which the value of these exponents can be derived such as the state of a pseudo-random generator used to compute these exponents~\cite{PFS}.

\section{Cryptographic properties}
To summarize the previous sections, the protocol should satisfy the following cryptographic properties.
\begin{itemize}
  \item Authenticity
  \item Accountability
  \item Confidentiality
  \item Integrity
  \item \scriptsize Availability
\end{itemize}

